Discovery and analysis of pleiotropic eQTLs

### 2023-09-01 doc/presentation_230120_internal_seminar/intern_sem_230120.tex

% Add section slide
\AtBeginSection[]
{
    \begin{frame}
        \frametitle{Table of Contents}
        \tableofcontents[currentsection]
    \end{frame}
}

@@

        \begin{itemize}
            \item Genome-wise association studies: Association test in the population of variants on common traits, eg. height or Crohn's disease
            \item Used for frequent variants and common quantitative or qualitative traits
            \item High-throughput genetics: 3000 unique loci for over thousands of disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
            \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories\footnote{Watanabe et al 2019. 10.1038/s41588-019-0481-0}
        \end{itemize}
%
%         \vfill

@@

%
        \begin{itemize}
            \item Correlation between variants. Which is the causal variant?
            \item Which genes are important for disease?
            \item In which tissues are these variants active?
        \end{itemize}
%
        \vfill
%
        Questions partially addressed by looking at eQTLs in different tissues

@@

        \begin{itemize}
            \item Colocalization between 293 GWAS and 127 eQTL studies
            \item 25\%\footnote{SNP.PP.H4$\geq$0.5} of 10,000 leading SNPs, colocalized with at least one eQTL
            \item 5,000 eQTLs and 5,000 target genes\footnote{SNP.PP.H4$\geq$0.5}
            \item 138e3 colocalized variants with probability $\geq$ 0.75
            \item Percentage of GWAS loci explained by eQTL: histogram
        \end{itemize}

@@

        \begin{itemize}
            \item A large set of GWAS variants annotated with putative target genes and tissues
            \item A list of pleiotropic eQTLs and regions
            \item Pleiotropic eQTL region include immune and cell cycle regulators
            \item Enrichment of 5' UTR and transcript (missense) variants
            \item Increased number of associated egenes and etissues
            \item A web site to access the data
        \end{itemize}

@@

        \begin{itemize}
            \item Verify causality of eQTL genes on GWAS
            \item Analyze pleiotropic loci in more detail
        \end{itemize}

@@

        \begin{itemize}
            \item L\'eopoldine Lecerf (M1)
            \item P Paul, P Rihet, M Michel, S Marquet, S Spicuglia
        \end{itemize}
%
        \vfill
%
        Funding
%
        \begin{itemize}
            \item Institut Cancer et Immunologie - Aix-Marseille Univ.
            \item Agence nationale de la recherche (ANR)
            \item GOLD
        \end{itemize}

>>>>

% Add section slide
\AtBeginSection[]
{
    \begin{frame}
        \frametitle{Table of Contents}
        \tableofcontents[currentsection]
    \end{frame}Large scale GWAS/eQTL colocalization analysis
}

@@

        \begin{itemize}
            \item Genome-wise association studies: Test of association of variants to traits in a population, eg. height or Crohn's disease
            \item Used for frequent variants and common quantitative or qualitative traits
            \item High-throughput genetics: 3000 unique loci for over thousands of disease/trait outcomes\footnote{Loos 2020. 10.1038/s41467-020-19653-5}
            \item Of the 1,707.0 Mb genome, 90.0\% is associated to different GWAS categories\footnote{Watanabe et al 2019. 10.1038/s41588-019-0481-0}
        \end{itemize}
%
%         \vfill

@@

%
        \begin{itemize}
            \item Which is the causal variant?
            \item Which genes are important for disease?
            \item In which tissues are these variants active?
        \end{itemize}
%
        \vfill
%
        Questions partially addressed by looking at eQTLs in different tissues

@@

        \begin{itemize}
            \item Colocalization between 293 GWAS and 127 eQTL studies
            \item 25\%\footnote{SNP.PP.H4$\geq$0.5} of 10,000 leading SNPs, colocalized with at least one eQTL
            \item 5,000 target genes\footnote{SNP.PP.H4$\geq$0.5}
            \item 138e3 colocalized variants with probability $\geq$ 0.75
            \item Percentage of GWAS loci explained by eQTL: histogram
        \end{itemize}

@@

        \begin{itemize}
            \item A large set of GWAS variants annotated with putative target genes and tissues
            \item A list of pleiotropic eQTLs and regions
            \item Pleiotropic eQTL region include immune and cell cycle regulators
            \item Increased number of associated egenes and etissues
            \item A web site to access the data
        \end{itemize}

@@

        \begin{itemize}
            \item Verify causality of eQTL genes on GWAS
            \item Analyze pleiotropic loci in more detail
            \item Compute colocalization of new GWAS datasets
        \end{itemize}

@@

        \begin{itemize}
            \item L\'eopoldine Lecerf (M1)
            \item P Paul, P Rihet, M Michel, S Marquet, S Spicuglia
        \end{itemize}
%
        \vfill
%
        Funding
%
        \begin{itemize}
            \item Institut Cancer et Immunologie - Aix-Marseille Univ.
            \item Agence nationale de la recherche (ANR)
        \end{itemize}
